---
title: "stage_3_1_predict_nis_lda"
author: "DBPR"
date: "2026-02-02"
output: html_document
---

In this code we will construct an holistic model aimed at testing whether traits can predict NIS/ non-NIS status.

Libraries
```{r}
library(tidyverse)
#library(MASS)
library(ggrepel)
library(pROC)
options(scipen = 999)
```
#1) LDA
##1.1) Data prep.
```{r}
# --- 1. DATA PREPARATION ---
# Clean data, define groups, and select only numeric predictors (Body size excluded)
lda_data_clean <- read.csv("./data/sp_level_data.csv") %>%
  
  # Select Predictors
  dplyr::select(
    species,
    group,
    "Commonness"=n_presence,
    "MaxN"=mean_MaxN,
    "Depth range"=depth_range, 
    "Min. depth"=min_depth, 
    "Temperature range"=sbt_range,
    `Corals`=corals,
    `Rocks`=rocks, 
    `Sand`=sand, 
    `Seagrass`=seagrass, 
    `Algae`=macroalgae, 
    `Artificial`=artificial, # Substrates
    `Silt`=silt,
    `Habitat breadth`=B_prime_std) %>%
  na.omit() # LDA cannot handle NAs

# Define Colors for consistency
custom_colors <- c("Natives - Red Sea" = "#007989", 
                   "Lessepsians - Red Sea"     = "#D9643E", 
                   "Lessepsians - Med. Sea"     = "#F8B954", 
                   "Natives - Med. Sea" = "#073956")
```
##2.2) LDA function
```{r}
perform_lda_analysis <- function(data, g1, g2, title_text, text_x_pos = 0.82, csv_filename = NULL) {
  
  # A. Filter Data
  sub_data <- data %>% filter(group %in% c(g1, g2))
  sub_data$group <- droplevels(factor(sub_data$group))
  
  # B. ROBUST FILTERING
  is_robust_trait <- function(x) {
    if(!is.numeric(x)) return(TRUE) 
    has_variance <- var(x, na.rm=TRUE) > 0
    is_common    <- sum(x > 0, na.rm=TRUE) >= 5 
    return(has_variance & is_common)
  }
  
  sub_data <- sub_data %>% select_if(is_robust_trait)
  
  # C. Run Standard LDA (For loadings and plotting)
  lda_model <- MASS::lda(group ~ ., data = sub_data %>% dplyr::select(-species))
  predictions <- predict(lda_model)
  
  # --- NEW STEP: Cross-Validated AUC (Replaces MANOVA) ---
  # Run LDA again with CV=TRUE to get unbiased, out-of-sample posterior probabilities
  lda_cv <- MASS::lda(group ~ ., data = sub_data %>% dplyr::select(-species), CV = TRUE)
  
  # Create a binary response variable (1 for g2, 0 for g1)
  actual_class <- ifelse(sub_data$group == g2, 1, 0)
  
  # Extract the cross-validated probabilities for group g2
  predicted_probs <- lda_cv$posterior[, as.character(g2)]
  
  # Calculate the Area Under the Curve (AUC)
  roc_obj <- roc(actual_class, predicted_probs, quiet = TRUE)
  cv_auc  <- as.numeric(auc(roc_obj))
  
  # Create the new, statistically sound plot label
  plot_label <- paste0("LOOCV AUC = ", round(cv_auc, 2))
  
  # --- CREATE SUMMARY DATAFRAME ---
  stats_df <- data.frame(
    Comparison = paste(g1, "vs", g2),
    LOOCV_AUC = round(cv_auc, 3)
  )
  #EXPORT CV PROBABILITIES TO CSV
  cv_probs_df <- data.frame(
    species = sub_data$species,
    lda_cv_prob = predicted_probs
  )
  
  # Clean the title_text to make a safe filename (removes spaces)
  safe_title <- gsub(" ", "_", title_text)
  cv_filename <- paste0("./models/", safe_title, "_cv_probs.csv")
  write.csv(cv_probs_df, cv_filename, row.names = FALSE)
  message(paste("Saved CV probabilities to:", cv_filename))
  
  # E. Extract Loadings
  loadings <- as.data.frame(lda_model$scaling)
  colnames(loadings) <- "LD1"
  loadings$Trait <- rownames(loadings)
  loadings$Trait <- gsub("`", "", rownames(loadings))
  loadings <- loadings %>% arrange(desc(abs(LD1)))
  
  # EXPORT TO CSV
  if (!is.null(csv_filename)) {
    write.csv(loadings, file = csv_filename, row.names = FALSE)
    message(paste("Saved loadings to:", csv_filename))
  }
  
  # F. Plotting
  plot_data <- data.frame(
    species = sub_data$species,
    Group = sub_data$group,
    LD1_Score = predictions$x[,1]
  )
  
  # Scaling Arrows
  max_score <- max(abs(plot_data$LD1_Score))
  max_loading <- max(abs(loadings$LD1))
  arrow_scale <- (max_score / max_loading) * 0.8 
  
  top_traits <- loadings %>% head(5)
  
  # DYNAMIC X POSITION
  x_range <- max(plot_data$LD1_Score) - min(plot_data$LD1_Score)
  x_pos   <- min(plot_data$LD1_Score) + (text_x_pos * x_range)

  p <- ggplot() +
    geom_density(data = plot_data, aes(x = LD1_Score, fill = Group), 
                 alpha = 0.4, color = NA) +
    geom_rug(data = plot_data, aes(x = LD1_Score, color = Group)) +
    
    geom_segment(data = top_traits, aes(x = 0, y = 0, xend = LD1 * arrow_scale, yend = 0),
                 arrow = arrow(length = unit(0.2, "cm")), color = "black", linewidth = 0.8) +
    
    geom_text_repel(data = top_traits, aes(x = LD1 * arrow_scale, y = 0, label = Trait),
                    direction = "y", nudge_y = 0.1, fontface = "bold", size = 3, 
                    max.overlaps = 15, box.padding = .65) +
    
    #Now displays the CV-AUC
    annotate("text", x = x_pos, y = Inf, label = plot_label, 
             hjust = 0, vjust = 1.3, size = 3, lineheight = 0.9) +
    
    scale_fill_manual(name = "Assemblage", values = custom_colors) +
    scale_color_manual(name = "Assemblage", values = custom_colors) +
    
    theme_bw() +
    theme(
      panel.grid = element_blank(),
      legend.position = c(0.02, 0.98), 
      legend.justification = c(0, 1),
      legend.background = element_rect(fill = "NA", color = "NA", linewidth = 0.5),
      axis.title = element_blank()
    ) +
    
    labs(title = title_text,
         x = "LD 1", y = "Density")
  
  return(list(
    model = lda_model, 
    loadings = loadings, 
    plot = p, 
    stats_table = stats_df,
    cv_auc = cv_auc # Exposing the AUC score directly for  ensemble step later!
  ))
}
```

##2.3) LDA analyses
```{r}
# 1) Red Sea Natives vs Red Sea NIS (Invasiveness)
# "What makes a species capable of leaving?"
lda_1 <-
  perform_lda_analysis(
    lda_data_clean,
    "Natives - Red Sea",
    "Lessepsians - Red Sea",
    "A) Native range comparison",
    csv_filename = "./models/native_range_comp_loadings.csv"
  )
lda_1
saveRDS(lda_1$model,"./models/lda_1.rds")
# 2) Red Sea NIS vs Med Sea NIS (filtering)
# "Did they change during transit?"
# FIX: Removing Silt to solve VIF collinearity with Sand and Algae
lda_2 <-
  perform_lda_analysis(
    lda_data_clean %>% dplyr::select(-Silt), 
    "Lessepsians - Red Sea",
    "Lessepsians - Med. Sea",
    "B) Native-invaded range comparison",
    csv_filename = "./models/native_invaded_range_comp_loadings.csv"
  )
lda_2

# 3) Med Sea NIS vs Med Sea Natives (establishment)
# "How do they fit into the new community?"
# FIX: Removing Rocks (previous fix) and Silt (to solve VIF collinearity with Min. depth)
lda_3 <-
  perform_lda_analysis(
    lda_data_clean %>% dplyr::select(-Rocks, -Silt),
    "Lessepsians - Med. Sea",
    "Natives - Med. Sea",
    "C) Invaded range comparison",
    csv_filename = "./models/invaded_range_comp_loadings.csv"
  )
lda_3
```

##2.4) Sensitivity analysis: Shared habitats only
```{r}
# --- SENSITIVITY: Test Native-Invaded shift ONLY on shared environmental space ---
# We might argue shifts are obvious due to lack of corals/different temps in the Med.
# We strip away variables that are completely missing (zero variance) in ONE of the basins.

# 1. Filter for just the Lessepsians across both basins
sens_data <- lda_data_clean %>% 
  filter(group %in% c("Lessepsians - Red Sea", "Lessepsians - Med. Sea"))

# 2. Identify variables that are completely absent (sum = 0) in EITHER basin
# By checking group variance, we find variables like 'Corals' which are 0 in the Med.
valid_vars <- names(sens_data)[sapply(names(sens_data), function(var) {
  if(!is.numeric(sens_data[[var]])) return(TRUE) # keep species/group
  
  # Check sum in Red Sea
  sum_rs <- sum(sens_data[[var]][sens_data$group == "Lessepsians - Red Sea"], na.rm = TRUE)
  # Check sum in Med Sea
  sum_med <- sum(sens_data[[var]][sens_data$group == "Lessepsians - Med. Sea"], na.rm = TRUE)
  
  # Keep only if present in BOTH (sum > 0)
  return(sum_rs > 0 & sum_med > 0)
})]

# We also manually exclude Temperature Range as the Med is inherently more variable seasonally
sens_data_filtered <- sens_data %>% 
  dplyr::select(all_of(valid_vars)) %>%
  dplyr::select(-`Temperature range`,-Seagrass)

print("Variables retained for Sensitivity Analysis (Shared Space Only):")
print(names(sens_data_filtered))

# 3. Run the LDA on this restricted space
lda_2_sens <- perform_lda_analysis(
    sens_data_filtered,
    "Lessepsians - Red Sea",
    "Lessepsians - Med. Sea",
    "Sensitivity: Shared Space Comparison",
    csv_filename = "./models/native_invaded_sensitivity_loadings.csv"
  )
lda_2_sens
print("--- SENSITIVITY LDA RESULTS ---")
print(lda_2_sens$stats_table)
# If this is still significant, it proves active behavioral flexibility (Niche Release)
# rather than just passive displacement caused by missing habitats.
```

##2.5) Plotting
```{r}
# --- 4. COMBINED PLOT ---
final_plot <- gridExtra::arrangeGrob(
  lda_1$plot,
  lda_2_sens$plot+labs(title = "B) Native-invaded range comparison"),
  lda_3$plot,
  ncol = 1,
  bottom = "LD 1",
  left = "Density"
)
ggsave("./plots/Figure_3.png", final_plot, width = 15, height = 25, units = "cm", dpi = 300,bg = "white")
ggsave("./plots/Figure_3.pdf", final_plot, width = 15, height = 25, units = "cm", dpi = 300,bg = "white",device = cairo_pdf)
```

Plot lda 2
```{r}
ggsave(
  "./plots/SI_lda_2.png",
  lda_2$plot +
    theme(title = element_blank()),
  width = 15,
  height = 10,
  units = "cm",
  dpi = 300,
  bg = "white",
  device = "png"
)

```

#3. Predicting future invaders
```{r}
# A. Determine the "Direction of Invasiveness" using data stored in the plot object
# We access the dataframe used for the density plot (lda_1$plot$data)
# This guarantees the Scores and Groups are aligned.
plot_data <- lda_1$plot$layers[[1]]$data

train_means <- tapply(plot_data$LD1_Score, plot_data$Group, mean)
# B. Check direction
# If mean(Lessepsians) < mean(Natives), we flip the sign (-1)
# High Score must equal High Risk
nis_mean    <- train_means["Lessepsians - Red Sea"]
native_mean <- train_means["Natives - Red Sea"]
direction   <- ifelse(nis_mean > native_mean, 1, -1)

# C. Project Red Sea Natives
rs_natives <- lda_data_clean %>% filter(group == "Natives - Red Sea")

# Predict
pred_invaders <- predict(lda_1$model, newdata = rs_natives %>% dplyr::select(-species, -group))

# D. Assign and Correct Scores
rs_natives$Invasion_Potential <- pred_invaders$x[,1] * direction

# E. Sort and Save (Body size excluded from select)
ranked_candidates <- rs_natives %>% 
  arrange(desc(Invasion_Potential)) %>% 
  dplyr::select(species, Invasion_Potential, Algae, MaxN, `Depth range`, Corals, Rocks, Seagrass, Artificial, Silt)

# Save
write.csv(ranked_candidates, "./models/watchlist_lda.csv", row.names = FALSE)

print(paste0("Direction Correction applied: multiplied scores by ", direction))
print("--- TOP 10 RED SEA NATIVES MOST LIKELY TO INVADE ---")
head(ranked_candidates, 10)

#save posterior probs.
# 1. Explicitly predict ONLY on the 108 native species
# This forces the output to be exactly 108 rows long
lda_1_preds <- predict(lda_1$model, newdata = rs_natives)

# 2. Create the standalone data frame
lda_1_posterior_output_df <- data.frame(
  species = rs_natives$species,
  probs = lda_1_preds$posterior[, "Lessepsians - Red Sea"]
)

# 3. (Optional but recommended) Sort the list from highest risk to lowest risk
lda_1_posterior_output_df <- lda_1_posterior_output_df %>%
  arrange(desc(probs))

write.csv(lda_1_posterior_output_df,"./models/lda_1_posterior_output_df.csv",row.names = F)
```

Export loadings table
```{r}
# 1. Read the individual results
loadings_1 <- read.csv("./models/native_range_comp_loadings.csv") %>% 
  rename(`Native range comparison` = LD1)

loadings_2 <- read.csv("./models/native_invaded_range_comp_loadings.csv") %>% 
  rename(`Native-invaded range comparison` = LD1)

loadings_3 <- read.csv("./models/invaded_range_comp_loadings.csv") %>% 
  rename(`Invaded range comparison` = LD1)

# Read the new sensitivity analysis results
loadings_sens <- read.csv("./models/native_invaded_sensitivity_loadings.csv") %>% 
  rename(`Native-invaded (Sensitivity)` = LD1)

# 2. Merge them into one table
all_loadings <- loadings_1 %>%
  full_join(loadings_2, by = "Trait") %>%
  full_join(loadings_sens, by = "Trait") %>% # Add the sensitivity column
  full_join(loadings_3, by = "Trait") %>%
  arrange(Trait) %>%  # Sort alphabetically by trait name
  dplyr::select(
    Trait,
    `Native range comparison`,
    `Native-invaded range comparison`,
    `Native-invaded (Sensitivity)`,
    `Invaded range comparison`
  )

# 3. Export
write.csv(all_loadings, "./models/lda_loadings_summary.csv", row.names = FALSE)

# 4. Print for verification
print(all_loadings)
```

##2.4) LDA assumptions
```{r}
##2.6) LDA assumptions
library(biotools)
library(car)

check_lda_assumptions <- function(data, g1, g2) {
  
  # 1. Prep Data (Same cleaning as before)
  sub_data <- data %>% filter(group %in% c(g1, g2))
  sub_data$group <- factor(sub_data$group)
  
  # Remove Low Variance Columns (Same logic as robust LDA)
  is_robust <- function(x) {
    if(!is.numeric(x)) return(TRUE)
    var(x, na.rm=TRUE) > 0 && sum(x > 0, na.rm=TRUE) >= 5
  }
  sub_data <- sub_data %>% select_if(is_robust)
  
  # Predictors Matrix
  predictors <- sub_data %>% dplyr::select(-species, -group)
  
  cat(paste0("\n--- ASSUMPTION CHECK: ", g1, " vs ", g2, " ---\n"))
  
  # --- CHECK 1: MULTICOLLINEARITY ---
  # If variables are highly correlated, loadings (rankings) become unstable.
  # We check Variance Inflation Factor (VIF). VIF > 10 is bad.
  cat("\n1. Multicollinearity (Max VIF):\n")
  tryCatch({
    # We fit a dummy linear model just to get VIF values
    dummy_mod <- lm(as.numeric(sub_data$group) ~ ., data = predictors)
    vif_vals <- car::vif(dummy_mod)
    print(vif_vals[vif_vals > 5]) # Only show problematic ones
    if(max(vif_vals) < 5) cat("  - Pass: All VIFs < 5 (Low correlation)\n")
  }, error = function(e) cat("  - Could not calc VIF (Matrix singular)\n"))

  # --- CHECK 2: HOMOGENEITY OF COVARIANCES (Box's M) ---
  # LDA assumes both groups have the same "shape" (variance).
  cat("\n2. Homogeneity of Covariances (Box's M):\n")
  tryCatch({
    box_res <- biotools::boxM(predictors, sub_data$group)
    print(box_res)
    if(box_res$p.value < 0.05) {
      cat("  - WARNING: Assumption Violated. Groups have different variances.\n")
      cat("  - Note: This is expected with unbalanced sample sizes (144 vs 22).\n")
    } else {
      cat("  - Pass: Variances are homogeneous.\n")
    }
  }, error = function(e) cat("  - Box's M failed (likely too few NIS samples).\n"))
  
  # --- CHECK 3: NORMALITY (Visual) ---
  # Ecological data (like coral %) is rarely normal. 
  # We check if the LD scores look roughly bell-shaped.
  cat("\n3. Normality Check:\n")
  cat("  - See generated QQ-Plots. If points deviate from the line, data is non-normal.\n")
  
  # Run temp LDA for plotting
  lda_mod <- MASS::lda(group ~ ., data = sub_data %>% dplyr::select(-species))
  scores <- predict(lda_mod)$x[,1]
  
  qqnorm(scores, main = paste("QQ Plot: ", g1, "vs", g2)); qqline(scores, col = "red")
}

# RUN THE CHECKS
par(mfrow=c(1,3)) # View 3 plots at once

# 1. Native range (No collinearity issues, keep as is)
check_lda_assumptions(lda_data_clean, "Natives - Red Sea", "Lessepsians - Red Sea")

# 2. Native-invaded range (Remove Silt)
check_lda_assumptions(lda_data_clean %>% dplyr::select(-Silt), "Lessepsians - Red Sea", "Lessepsians - Med. Sea")

# 3. Invaded range (Remove Rocks and Silt)
check_lda_assumptions(lda_data_clean %>% dplyr::select(-Rocks, -Silt), "Lessepsians - Med. Sea", "Natives - Med. Sea")
```

Summary:
We can trust the Loadings: Since VIFs are good, we can confidently say which traits matter (e.g., "Corals" and "Macroalgae").

We can trust the LD1 Axis: The general direction of separation is likely real.

We should distrust the P-value (Slightly): If p=0.04, be skeptical. If p=0.005 (like we found), it's safe even with the violation.

Extract species scores
```{r}
lda_1$plot$layers[[1]]$data %>%
  arrange(desc(LD1_Score))

lda_2$plot$layers[[1]]$data %>%
  arrange(desc(LD1_Score))

lda_2_sens$plot$layers[[1]]$data %>%
  arrange(desc(LD1_Score))

lda_3$plot$layers[[1]]$data %>%
  arrange(desc(LD1_Score))
```

LDAs summari table
```{r}
# Combine the stats tables from all four LDA models
lda_summary_table <- dplyr::bind_rows(
  lda_1$stats_table,
  lda_2$stats_table,
  lda_3$stats_table,
  lda_2_sens$stats_table
)

# Overwrite the Comparison column with the standardized manuscript terminology
lda_summary_table$Comparison <- c(
  "Native range comparison",
  "Native-invaded range comparison",
  "Invaded range comparison",
  "Native-invaded range comparison (sensitivity)"
)
lda_summary_table
write.csv(lda_summary_table,"./models/lda_summary_table.csv",row.names = F)
```

