---
title: "stage_3_2_predict_nis_glmm"
author: "DBPR"
date: "2026-02-02"
output: html_document
---

In this code we will construct an holistic model aimed at testing whether traits can predict invaders/ non-invaders status.

Libraries
```{r}
library(tidyverse)
library(lme4)
library(DHARMa)
library(performance)
library(broom.mixed)
library(gridExtra)
library(car)      # For calculating VIF
library(pbapply)  # For a progress bar
library(glmm.hp)
options(scipen = 999)
```

#1) Setup and data prep.

```{r}
# --- MASTER DATA PREPARATION ---
data_full <- read.csv("./data/sp_level_data.csv") %>% 
  # Ensure random effects are factors
  mutate(Family = as.factor(Family))

# A. Filter & Prep
data_m1 <- data_full %>%
  filter(group %in% c("Natives - Red Sea", "Lessepsians - Red Sea")) %>%
  mutate(
    # Create Binary Response: Natives=0, NIS=1
    response = ifelse(group == "Lessepsians - Red Sea", 1, 0),
    group_label = droplevels(factor(group)),
    Family = droplevels(Family)
  )
```

#2) Dredge 
all possible models and find the best model
```{r warning=FALSE}
library(MuMIn)

# IMPORTANT: MuMIn requires this setting to run dredge. 
# It ensures the dataset doesn't change size due to missing values.
options(na.action = "na.fail")

# 1. Fit the Global Model
# This is the maximal model containing all 16 predictors and the random effect
cat("Fitting the global model...\n")
global_model <- glmer(
  response ~ 
    scale(mean_MaxN) + 
    scale(n_presence) + 
    scale(depth_range) + 
    scale(min_depth) + 
    scale(sbt_min) + 
    scale(sbt_range) + 
    scale(sbt_max) + 
    scale(artificial) + 
    scale(corals) + 
    scale(seagrass) + 
    scale(silt) + 
    scale(gravel) + 
    scale(sand) + 
    scale(rocks) + 
    scale(B_prime_std) + 
    scale(macroalgae) + 
    (1 | Family),
  data = data_m1,
  family = binomial
)

# 2. Run the Dredge Function
# 2. Run or Load the Dredge Function
model_file_path <- "./models/dredge/all_models.rds"

if (file.exists(model_file_path)) {
  
  cat("Saved dredge file found! Loading models instantly to save time...\n")
  all_models <- readRDS(model_file_path)
  
} else {
  
  cat("No saved file found. Starting dredge (this may take several hours)...\n")
  # Run the heavy computation
  all_models <- dredge(global_model)
  
  # Immediately save it so never have to run it again
  saveRDS(all_models, model_file_path)
  write.csv(all_models %>% as_tibble(), "./junk/all_models.csv", row.names = FALSE)
  cat("Dredge complete and saved!\n")
  
}

# 3. View the Results
# Look at the top 10 models ranked by AICc (dredge uses AICc by default for small sample sizes)
cat("\nTop models ranked by AICc:\n")
head(all_models, 10)

# 4. Extract the Best Model
# Get the absolute best model (rank 1)
best_dredged_model <- get.models(all_models, 1)[[1]]

cat("\nSummary of the BEST dredged model:\n")
summary(best_dredged_model)

# Reset na.action to default after are done
options(na.action = "na.omit")
```

relative importance of each predictor
```{r}
all_models=read_rds("./models/dredge/all_models.rds")
importance_ranking <- MuMIn::sw(all_models)

# 1. Extract the data directly from the MuMIn 'importance_ranking' object
importance_df <- data.frame(
  Predictor = names(importance_ranking),
  Sum_of_Weights = as.numeric(importance_ranking),
  N_Models = attr(importance_ranking, "n.models")
)

# 2. Clean up the predictor names (Remove "scale(" and ")" for a polished look)
importance_df$Predictor <- gsub("scale\\(|\\)", "", importance_df$Predictor)

# 3. write result
write.csv(importance_df,"./junk/importance_df.csv",row.names = F)
```

#3) Model 1: Native range comparison Comparison:
Red Sea natives (0) vs. Red Sea NIS (1) Hypothesis: Who leaves?
```{r}
# B. Run the best model
# Removed body size from the formula
m1 <- glmer(
  response ~ 
    scale(min_depth) + 
    scale(corals) +
    (1 | Family),
  data = data_m1,
  family = binomial)
#saveRDS(m1,"./models/native_r_comp_glmm.rds")
# C. Diagnostics
summary(m1)
MuMIn::AICc(m1)
performance::check_collinearity(m1)
performance::check_convergence(m1)
performance::check_singularity(m1)
r2_m1 <- performance::r2_nakagawa(m1)
print(r2_m1)
# DHARMa Residuals
plot(simulateResiduals(m1))
```

##2.1) summary
```{r}
m1_res <- broom.mixed::tidy(m1, conf.int = TRUE, conf.level = 0.95) %>%
  filter(effect == "fixed", term != "(Intercept)") %>%
  mutate(
    term_clean = case_when(
      grepl("min_depth", term) ~ "Min. depth",
      grepl("corals", term) ~ "Coral association",
      TRUE ~ term
    ),
    is_significant = ifelse(p.value < 0.05, "Significant", "Not Significant")
  ) %>% 
  select(-group,-effect,-term) %>% 
  select(term_clean, everything())

r2_label_m1 <- paste0("R²(m)=", round(r2_m1$R2_marginal,2), " | R²(c)=", round(r2_m1$R2_conditional,2))
```

#3) Predict future invaders
Here we apply the model trained in Step 2 to the current Red Sea native community. We identify which native species have the highest probability of being “misclassified” as invaders by the model (i.e., they have high invasion potential based on their traits and phylogeny).
```{r}
# 1. Filter for Red Sea Natives only (The candidates)
# We ensure we drop NAs so the rows align with predictions
# Removed body size from the select statement
candidates <- data_full %>% 
  filter(group == "Natives - Red Sea") %>%
  select(species, Family, min_depth, corals) %>%
  na.omit()

# 2. Predict Probability
# type = "response" gives us the probability scale (0 to 1)
candidates$invasion_prob <- predict(m1, newdata = candidates, type = "response")

# 3. Create Ranked List
watchlist <- candidates %>%
  select(species, Family, invasion_prob, corals) %>%
  arrange(desc(invasion_prob)) %>%
  mutate(
    Rank = row_number(),
    Probability = round(invasion_prob, 3)
  ) %>%
  select(Rank, species, Family, Probability, everything(), -invasion_prob)

# 4. Display Top 20 Candidates
print("--- TOP 20 NATIVE SPECIES WITH HIGHEST INVASION POTENTIAL ---")
head(watchlist, 20)
write.csv(watchlist, "./models/watchlist_glmm_best_model.csv", row.names = FALSE)
```

#4) Variance partition
```{r}
#Prepare model for glmm.hp
data_m1_var_p=data_m1 %>% 
  mutate(
         s_min_depth=scale(min_depth),
         s_corals=scale(corals))

#Run GLMM with pre-scaled variables
m1_var_p <- lme4::glmer(
  response ~ 
    s_min_depth + 
    s_corals +
    (1 | Family),
  data = data_m1_var_p,
  family = binomial)

library(glmm.hp)
R2_partition_results <-glmm.hp(m1_var_p,
                                type="R2")

R2_partition_results
```

#5) Plot
```{r}
# Extract variance partition results from glmm.hp
# The theoretical table contains the 'Individual' and 'I.perc(%)' columns
vp_df <- as.data.frame(R2_partition_results$theoretical)
vp_df$term <- rownames(vp_df)

# 2. Clean terms to exactly match  m1_res labels
vp_df <- vp_df %>%
  mutate(
    term_clean = case_when(
      grepl("s_min_depth", term) ~ "Min. depth",
      grepl("s_corals", term) ~ "Coral association",
      TRUE ~ term
    )
  )

# Join with tidied GLMM results and create a rich axis label
m1_res <- m1_res %>%
  left_join(vp_df, by = "term_clean") %>%
  mutate(
    # Create a label like: "Coral association (R² = 0.252, 62.9%)"
    # Using sprintf ensures 3 decimal places are always shown for neatness
    plot_label = paste0(term_clean, " (R²m = ", sprintf("%.3f", Individual), ", ", round(`I.perc(%)`, 1), "%)")
  )

# Generate the plot
p1 <-
  ggplot(m1_res, aes(
    x = estimate,
    y = reorder(plot_label, estimate), # Reorder using the new rich label
    color = is_significant
  )) +
  geom_vline(
    xintercept = 0,
    linetype = "dashed",
    color = "grey50",
    size = 0.8
  ) +
  geom_pointrange(aes(xmin = conf.low, xmax = conf.high),
                  size = 0.6,
                  flatten = 0.5) +
  scale_color_manual(values = c(
    "Not Significant" = "grey70",
    "Significant" = "#D9643E"
  )) +
  theme_bw() +
  theme(
    legend.position = "none", 
    panel.grid.minor = element_blank(),
    axis.text.y = element_text(size = 10) # Ensures the longer labels are readable
  ) +
  labs(
    x = "Standardized coefficient",
    y = "",
    caption = r2_label_m1
  )
p1
ggsave("./plots/Figure_4_best_model.png", p1, 
       width = 15, height = 15, units = "cm", dpi = 300)
```
