---
title: "stage_2_trait_space"
author: "Shahar Chaikin"
date: "2026-02-03"
output: html_document
---

Produce holistic ordination aimed at testing the level of separation across assemblages

Libraries
```{r}
library(FD)
library(vegan)
library(tidyverse)
library(ggrepel)
library(gridExtra)
library(pairwiseAdonis)
```

#1) Prepare data
```{r}
sp_level_data <- read.csv("./data/sp_level_data.csv") %>%
  
  # C. Create Unique Labels for Plotting
  # This ensures "Siganus rivulatus" appears as "S. rivulatus (RS)" and "S. rivulatus (Med)"
  mutate(
    plot_label = paste0(species, " (", ifelse(sea=="red", "RS", "Med"), ")")
  )
#Traits data (Body size excluded)
traits <- sp_level_data %>%
  select(`MaxN`=mean_MaxN,
         `Commonness`=n_presence, 
         `Depth range`=depth_range,
         `Min. depth`=min_depth,
         `Min. SBT`=sbt_min,
         `SBT range`=sbt_range,
         `Max. SBT`=sbt_max,
         `Artificial`=artificial, 
         `Corals`=corals,
         `Seagrass`=seagrass,
         `Silt`=silt,
         `Gravel`=gravel,
         `Sand`=sand,
         `Rocks`=rocks,
         `Habitat breadth`=B_prime_std,
         `Algae`=macroalgae)
```

#2) Gower Distance & PCoA
```{r}
dist_matrix <- gowdis(traits)
pcoa_res <- pcoa(dist_matrix)
# Extract specifically the "Relative_eig" column (Variance Explained)
# This gives the proportion (0 to 1) for every axis
variance_explained <- pcoa_res$values$Relative_eig

# Print the percentage for the first two axes (the ones plotted)
axis_1_pct <- round(variance_explained[1] * 100, 1)
axis_2_pct <- round(variance_explained[2] * 100, 1)

print(paste0("PCoA Axis 1 explains: ", axis_1_pct, "%"))
print(paste0("PCoA Axis 2 explains: ", axis_2_pct, "%"))
# Extract Scores
pcoa_df <- as.data.frame(pcoa_res$vectors[, 1:2])
colnames(pcoa_df) <- c("PCoA1", "PCoA2")
# Bind metadata safely
pcoa_df <- cbind(pcoa_df, sp_level_data %>% select(species, group, plot_label, sea, status))
write.csv(pcoa_df,"./models/pcoa_df.csv",row.names = F)
# Calculate Centroids (for the Diamonds)
centroids <- pcoa_df %>%
  group_by(group) %>%
  summarise(PCoA1 = mean(PCoA1), PCoA2 = mean(PCoA2))

# A) Test for Homogeneity of Multivariate Dispersions
# This checks if the groups have different "spreads" (variances)
disp_test <- betadisper(dist_matrix, sp_level_data$group)
print(anova(disp_test)) 
# If p > 0.05, groups have similar spread (ideal). 
# If p < 0.05, one group occupies a larger volume of trait space.
# Test exactly which groups differ (Tukey's Honest Significant Difference)
# This tells if the "Lessepsians - Med. Sea" are significantly more variable than "Lessepsians - Red Sea"
tukey_disp <- TukeyHSD(disp_test)
print(tukey_disp)

# Calculate correlation between traits and PCoA axes (Loadings)
trait_loadings <- envfit(pcoa_res$vectors[,1:2], traits, permutations = 999)
```

Export loadings
```{r}
#Extract the scores (direction vectors)
loadings_scores <- as.data.frame(scores(trait_loadings, display = "vectors"))

#Extract the statistics (R-squared and P-values) from the envfit object
loadings_summary <- data.frame(
  Trait = rownames(loadings_scores),
  PCoA1 = loadings_scores$Axis.1,
  PCoA2 = loadings_scores$Axis.2,
  r2 = trait_loadings$vectors$r,
  p_value = trait_loadings$vectors$pvals
) %>%
  arrange(p_value, desc(r2)) # Sort by significance (p) then strength (r2)

#Save
write.csv(loadings_summary, "./models/pcoa_trait_loadings.csv", row.names = FALSE)

#View top drivers
print(head(loadings_summary))
```

Export dispersion
```{r}
# 1. Convert Tukey output to a dataframe
tukey_df <- as.data.frame(tukey_disp$group)

# 2. Add the comparison names as a proper column (currently they are row names)
tukey_df$Comparison <- rownames(tukey_df)

# 3. Reorder columns for readability and round values
tukey_summary <- tukey_df %>%
  select(Comparison, diff, lwr, upr, p_adj = `p adj`) %>%
  mutate(
    diff = round(diff, 4),
    lwr = round(lwr, 4),
    upr = round(upr, 4),
    p_adj = round(p_adj, 4),
    Significance = ifelse(p_adj < 0.05, "*", "ns") # Add significance stars
  ) %>%
  arrange(p_adj) # Sort by most significant

# 4. Save to CSV
write.csv(tukey_summary, "./models/betadisper_tukey_results.csv", row.names = FALSE)

# Print for verification
print(tukey_summary)
```

#3) partitioning distance matrices among assemblages
```{r}
global_permanova <- adonis2(dist_matrix ~ group, data = sp_level_data, permutations = 9999)
print(global_permanova)
# B) Pairwise PERMANOVA
pairwise_res <- pairwise.adonis(dist_matrix, sp_level_data$group,perm = 9999)
print(pairwise_res)
write.csv(pairwise_res,"./models/pairwise_permanova__res.csv",row.names = F)
# 1. Visualize the Niche Breadth (Distance to Centroid)
# This boxplot will show  who has the "biggest" niche
boxplot(disp_test, xlab = "Assemblage", ylab = "Distance to centroid")
```

#4) Area overlaps
Calculate overlaps and area relationships
```{r}
library(sf)
# --- 1. DATA PREPARATION ---
hull_data <- pcoa_df %>% 
  ungroup() %>%
  mutate(group_factor = factor(group, 
                               levels = c("Natives - Red Sea", 
                                          "Lessepsians - Red Sea", 
                                          "Lessepsians - Med. Sea", 
                                          "Natives - Med. Sea")))

# --- 2. GENERATE HULLS (The Base R Way) ---
# We use chull() to find points and store them in a standard dataframe
hulls_df <- data.frame()

for(g in levels(hull_data$group_factor)) {
  sub_dat <- hull_data %>% filter(group_factor == g)
  
  # Calculate hull indices
  if(nrow(sub_dat) >= 3) {
    hull_idx <- chull(sub_dat$PCoA1, sub_dat$PCoA2)
    
    # Extract points and close the loop (first point = last point)
    hull_pts <- sub_dat[c(hull_idx, hull_idx[1]), ]
    
    # Add to master dataframe
    hulls_df <- rbind(hulls_df, hull_pts)
  }
}

# --- 3. BACKGROUND STATS (Using SF for intersection math only) ---
overlap_results <- data.frame()
# Helper to make SF polygon from points
make_poly <- function(df) {
  mat <- as.matrix(df[, c("PCoA1", "PCoA2")])
  st_polygon(list(mat)) %>% st_sfc() %>% st_as_sf()
}

# Loop for overlaps (Background calc)
groups <- levels(hull_data$group_factor)
pairs <- combn(groups, 2, simplify = FALSE)

for(p in pairs) {
  g1 <- p[1]; g2 <- p[2]
  df1 <- hulls_df %>% filter(group_factor == g1)
  df2 <- hulls_df %>% filter(group_factor == g2)
  
  if(nrow(df1) > 0 && nrow(df2) > 0) {
    poly1 <- make_poly(df1); poly2 <- make_poly(df2)
    a1 <- st_area(poly1); a2 <- st_area(poly2)
    inter <- st_intersection(poly1, poly2)
    a_shared <- if(length(inter) > 0) st_area(inter) else 0
    
    overlap_results <- rbind(overlap_results, data.frame(
      Group_A = g1, Group_B = g2,
      Overlap_Pct_of_A = round(as.numeric(a_shared/a1)*100, 1),
      Overlap_Pct_of_B = round(as.numeric(a_shared/a2)*100, 1)
    ))
  }
}

print(overlap_results) # stats table

# --- 4. CENTROIDS ---
centroids_all <- hull_data %>%
  group_by(group_factor) %>%
  summarise(PCoA1 = mean(PCoA1), PCoA2 = mean(PCoA2), .groups = "drop")
centroids_all
```

#5) Plot assemblages
```{r}
# --- 1. SETUP COLORS ---
custom_colors <- c("Natives - Red Sea" = "#007989", 
                   "Lessepsians - Red Sea"     = "#D9643E", 
                   "Natives - Med. Sea" = "#073956", 
                   "Lessepsians - Med. Sea"     = "#F8B954")

# --- 2. SPLIT DATA INTO LAYERS ---
# Instead of sorting, we just create two separate dataframes.
# This forces ggplot to draw them in the order we write the code.

# Background Layer (Natives)
hulls_natives <- hulls_df %>% filter(grepl("Natives", group_factor))
points_natives <- hull_data %>% filter(grepl("Natives", group_factor))

# Foreground Layer (NIS - Invaders)
hulls_nis <- hulls_df %>% filter(grepl("Lessepsians", group_factor))
points_nis <- hull_data %>% filter(grepl("Lessepsians", group_factor))

# --- 3. RECALCULATE ARROWS ---
trait_loadings <- envfit(pcoa_res$vectors[,1:2], traits, permutations = 999)
arrow_df <- as.data.frame(scores(trait_loadings, display = "vectors")) * 0.2
arrow_df$trait <- rownames(arrow_df)
arrow_df <- arrow_df %>%
  mutate(vector_len = sqrt(Axis.1^2 + Axis.2^2))

# --- 4. THE PLOT (TWO LAYERS) ---
plot_hulls_all <- ggplot() +
  
  # LAYER 1: Natives Polygons (Bottom)
  geom_polygon(data = hulls_natives, 
               aes(x = PCoA1, y = PCoA2, fill = group_factor), 
               alpha = 0.35, color = "black", size = 0.2) + 
  
  # LAYER 2: NIS Polygons (Top - Guaranteed)
  geom_polygon(data = hulls_nis, 
               aes(x = PCoA1, y = PCoA2, fill = group_factor), 
               alpha = 0.735, color = "black", size = 0.2) + 
  

  # LAYER 6: NIS Points (Top Points)
  geom_rug(
    data = points_nis,
    aes(x = PCoA1, y = PCoA2, color = group_factor),
    #size = 1.5,
    alpha = 0.6,
    #shape = 21,
    #color = "black",
    show.legend = F
  ) +
  
  # LAYER 5: Natives Points (Bottom Points)
  geom_rug(
    data = points_natives,
    aes(x = PCoA1, y = PCoA2, color = group_factor),
    #size = 1.5,
   # alpha = 0.6,
    #shape = 21,
    #color = "black",
    show.legend = F
  ) +
  
  # LAYER 3: Arrows
  geom_segment(data = arrow_df, 
               aes(x = 0, y = 0, xend = Axis.1, yend = Axis.2, alpha = vector_len),
               arrow = arrow(length = unit(0.2, "cm")), color = "black", size=0.6) +

  # LAYER 4: Trait Labels
  geom_text_repel(data = arrow_df, 
                  aes(x = Axis.1, y = Axis.2, label = trait, alpha = vector_len),
                  color = "black", fontface = "bold", size = 4.4,
                  box.padding = 0.8,
                  min.segment.length = 0
                  #bg.color="white", 
                  #bg.r=0.08
                  ) +
  # Scales & Legend
  scale_fill_manual(
    name = "Assemblage", 
    values = custom_colors,
    # THIS CONTROLS THE LEGEND ORDER:
    breaks = c("Natives - Red Sea", "Lessepsians - Red Sea", "Lessepsians - Med. Sea", "Natives - Med. Sea") 
  ) +
  scale_color_manual(
    name = "Assemblage", 
    values = custom_colors,
    # THIS CONTROLS THE LEGEND ORDER:
    breaks = c("Natives - Red Sea", "Lessepsians - Red Sea", "Lessepsians - Med. Sea", "Natives - Med. Sea") 
  )+
  scale_alpha_continuous(range = c(0.2, 1), guide = "none") +
  theme_classic() +
  labs(
    x = "PCoA 1",
    y = "PCoA 2",
    title = "A) Habitat space"#,
    # caption = paste0(
    #   "Native range comp.: P = 0.0001; ",
    #   "Native-invaded range comp.: P = 0.0001; ",
    #   "Invaded range comp.: P = 0.284"
  #  )
  )

print(plot_hulls_all)
```

#6) Shared assembalge
##6.1) Identify species present in BOTH seas

```{r}
shared_species_list <-  sp_level_data %>%
  group_by(species) %>%
  filter(n_distinct(sea) == 2) %>% # Must have exactly 2 sea entries
  pull(species) %>%
  unique()
```

##6.2) Extract PCoA Coordinates for these species from the Global PCoA
We use the Global Space so the axes (PCoA1, PCoA2) mean the same thing for both groups
```{r}
rs_coords <- pcoa_df %>% 
  filter(species %in% shared_species_list, sea == "red") %>%
  arrange(species) %>% # Sort alphabetically to ensure alignment
  select(species, PCoA1, PCoA2) %>%
  remove_rownames() %>%  
  column_to_rownames("species")

med_coords <- pcoa_df %>% 
  filter(species %in% shared_species_list, sea == "med") %>%
  arrange(species) %>% # Sort alphabetically to match RS
  select(species, PCoA1, PCoA2) %>%
    remove_rownames() %>%  
  column_to_rownames("species")

#Validation check
if(!identical(rownames(rs_coords), rownames(med_coords))) {
  stop("Error: Row names do not match! Check sorting.")
} else {
  print("Data aligned correctly. Ready for Procrustes.")
}
```

##6.3) Run Procrustes Analysis & Permutation Test
This step rotates the Red Sea configuration to best fit the Mediterranean configuration. The residuals represent the shift distance.
```{r}
# 1. Run Procrustes Analysis
# symmetric = TRUE calculates the fit without assuming one configuration is fixed
proc_res <- procrustes(X = rs_coords, Y = med_coords, symmetric = TRUE)

# 2. Test Significance (PROTEST)
# Tests if the similarity between RS and Med niches is greater than random
protest_res <- protest(X = rs_coords, Y = med_coords, permutations = 9999)

# Print Global Results
print(protest_res)
# Interpretation:
# - High Correlation (r): Niches are conserved (species do the same thing).
# - Low Correlation: High niche shift.
# - p-value (Is the match between these two maps better than random?): The RS configuration PREDICTS the Med configuration (non-random structure). In our case, non-significant means Species responded individually and chaotically


#A value of 0.3 is low for a Procrustes test. It means there is very little overlap or congruence between the two shapes. For context, "conserved" niches usually show correlations > 0.7 or 0.8.

#Sum of Squares represents the "mismatch" or error after trying to fit the two maps together. 0.91 (on a scale of 0 to 1) indicates that 81% of the variation in the Mediterranean position cannot be explained by the Red Sea position. This confirms that the shifts are massive.
```


##6.3) Extract Vector Shifts (Species-Specific Results) 
Here we calculate exactly how far each species moved in the trait space.
```{r}
# 1. Extract the shift vectors
# The 'residuals' from procrustes give the mismatch distance
shifts_df <- data.frame(
  species = rownames(rs_coords),
  RS_PCoA1 = rs_coords$PCoA1,
  RS_PCoA2 = rs_coords$PCoA2,
  Med_PCoA1 = med_coords$PCoA1,
  Med_PCoA2 = med_coords$PCoA2,
  total_shift = residuals(proc_res) # The Euclidean distance of the shift
)

#total_shift: This is the Euclidean distance (or Procrustes residual) between the Red Sea point and the Mediterranean point.
#High Value (> 0.3): Extreme Niche Shift (High Plasticity).
#Low Value (< 0.1): Niche Conservatism (Doing the same thing as back home).

# 2. Identify the Top 5 "Shifters" (Ecological Plasticity?)
top_shifters <- shifts_df %>%
  arrange(desc(total_shift)) %>%
  head(5)

print("Top 5 Species with Largest Niche Shifts:")
print(top_shifters[, c("species", "total_shift")])
```

Ranking Traits based on PCoA Shifts
This code requires the shifts_df (from Shift Plot) and arrow_df (from Environmental Drivers plot).
```{r}
# 1. Calculate the Species Movement Vector (The "Species Arrow")
# How much did the species move along Axis 1 and Axis 2?
species_vectors <- shifts_df %>%
  select(species, RS_PCoA1, RS_PCoA2, Med_PCoA1, Med_PCoA2) %>%
  mutate(
    delta_1 = Med_PCoA1 - RS_PCoA1, # Change in PCoA1
    delta_2 = Med_PCoA2 - RS_PCoA2  # Change in PCoA2
  ) %>%
  select(species, delta_1, delta_2)

# 2. Get Trait Vectors (The "Environment Arrows")
# We use arrow_df which contains the PCoA1/PCoA2 coordinates of each trait tip
trait_vectors <- arrow_df %>%
  select(trait, Axis.1, Axis.2)

# 3. Calculate Projections (The "Alignment")
# We loop through every species and every trait to see how well they align
projections <- expand.grid(species = species_vectors$species, 
                           trait = trait_vectors$trait) %>%
  left_join(species_vectors, by = "species") %>%
  left_join(trait_vectors, by = "trait") %>%
  # --- THE CORE CALCULATION (Dot Product) ---
  mutate(
    # How much of the species' shift is explained by this trait?
    projection = (delta_1 * Axis.1) + (delta_2 * Axis.2)
  ) %>%
  group_by(species) %>%
  # Normalize so we can compare (optional, but helps interpretation)
  arrange(species, desc(abs(projection)))

# 4. View the "Main Driver" for each species
top_contributors <- projections %>%
  slice(1) %>% # Get the #1 strongest driver per species
  select(species, trait, projection) %>%
  mutate(direction = ifelse(projection > 0, "Increased", "Decreased"))

print(top_contributors)
#Analysis of vector projections revealed two distinct drivers of niche shifts. One group of invaders (e.g., Pterois miles, Fistularia commersonii) was characterized by a shift away from Coral associations, reflecting a 'release' from reef-dependence in the coral-poor Mediterranean. A second group (e.g., Torquigener flavimaculosus, Siganus luridus) was defined by a shift into habitats with high Sea Bottom Temperature (SBT) ranges, indicating an expansion into more seasonally variable environments characteristic of the temperate recipient basin.
```

##6.4) Plot
Calculate the ratio and prints the conclusion based on Table 2 definitions
```{r}
library(sf)
# 1. Prepare Data: Select ALL Invaders (status == 1)
# We ensure we are using the full dataset, not just paired species
nis_all <- pcoa_df %>% 
  ungroup() %>%
  filter(status == 1) %>%
  mutate(sea_label = factor(sea, levels = c("red", "med"), 
                            labels = c("Native Range (Red Sea)", "Invaded Range (Med)")))

# 2. Calculate Hulls (Robust Function)
get_hull <- function(df) {
  if(nrow(df) < 3) return(NULL)
  coords <- df %>% select(PCoA1, PCoA2) %>% as.matrix()
  rownames(coords) <- NULL 
  hull_id <- chull(coords)
  poly_coords <- coords[c(hull_id, hull_id[1]), ]
  st_sf(geometry = st_sfc(st_polygon(list(poly_coords))))
}

rs_poly  <- get_hull(nis_all %>% filter(sea == "red"))
med_poly <- get_hull(nis_all %>% filter(sea == "med"))

# 3. Calculate Centroids (Mean X, Mean Y for each group)
centroids <- nis_all %>%
  group_by(sea) %>%
  summarise(
    PCoA1 = mean(PCoA1),
    PCoA2 = mean(PCoA2),
    .groups = "drop"
  ) %>%
  mutate(sea_label = factor(sea, levels = c("red", "med"), 
                            labels = c("Native Range (Red Sea)", "Invaded Range (Med)")))

# 4. Calculate Statistics for Caption
area_rs  <- st_area(rs_poly)
area_med <- st_area(med_poly)
ratio    <- as.numeric(area_med / area_rs)
intersection_poly <- st_intersection(rs_poly, med_poly)
overlap_pct <- as.numeric((st_area(intersection_poly) / area_med) * 100)
contraction_pct <- (1 - ratio) * 100
```

Plot shared
```{r}
# --- 1. PREPARE DATA (THE FIX) ---
# Instead of filter(status == 1), we keep anything labeled as "NIS" in the group column.
# This ensures Seriola and Echeneis are kept even if their original status was 0.
nis_all <- pcoa_df %>% 
  ungroup() %>%
  filter(grepl("Lessepsians", group)) %>% # <--- CHANGED FROM status == 1
  mutate(sea_label = factor(sea, levels = c("red", "med"), 
                            labels = c("Native Range (Red Sea)", "Invaded Range (Med)")))

# --- 2. CALCULATE HULLS ---
get_hull_df <- function(df, group_label) {
  coords <- df %>% select(PCoA1, PCoA2)
  hull_idx <- chull(coords$PCoA1, coords$PCoA2)
  hull_pts <- df[c(hull_idx, hull_idx[1]), ]
  return(hull_pts)
}

rs_hull_df <- get_hull_df(nis_all %>% filter(sea == "red"), "Native Range (Red Sea)")
med_hull_df <- get_hull_df(nis_all %>% filter(sea == "med"), "Invaded Range (Med)")

# --- 3. PREPARE ARROWS (Shared Species Only) ---
arrow_data <- nis_all %>%
  select(species, sea, PCoA1, PCoA2) %>%
  pivot_wider(names_from = sea, values_from = c(PCoA1, PCoA2)) %>%
  # Keep only species that have coordinates for BOTH seas
  filter(!is.na(PCoA1_red) & !is.na(PCoA1_med))

# --- 4. CENTROIDS ---
centroids <- nis_all %>%
  group_by(sea_label) %>%
  summarise(PCoA1 = mean(PCoA1), PCoA2 = mean(PCoA2), .groups = "drop")

# --- RE-DEFINE THE CAPTION 
stats_caption <- paste0("r = ", round(protest_res$t0, 3),
                        "; P = ", round(protest_res$signif, 3),
                        "; m² = ", round(protest_res$ss, 3))

# --- 5. THE PLOT ---
shared_sp_plot <- ggplot() +
  
  # A. Hulls
  geom_polygon(data = rs_hull_df, aes(x = PCoA1, y = PCoA2), 
               fill = "#D9643E", alpha = 0.735, color = "black", size = 0.2) +
  geom_polygon(data = med_hull_df, aes(x = PCoA1, y = PCoA2), 
               fill = "#F8B954", alpha = 0.735, color = "black", size = 0.2) +
  # D. Species Points
  geom_point(data = nis_all, aes(x = PCoA1, y = PCoA2, color = sea_label), 
             size = 2, alpha = 0.8,show.legend = F) +
  # B. Shift Arrows
  geom_segment(
    data = arrow_data,
    aes(
      x = PCoA1_red,
      y = PCoA2_red,
      xend = PCoA1_med,
      yend = PCoA2_med
    ),
    arrow = arrow(length = unit(0.2, "cm"), type = "open"),
    color = "black",
    size = 0.4
  ) +
  
  # C. Labels (Only Shared Species)
  geom_text_repel(data = arrow_data,
    aes(x = PCoA1_red, y = PCoA2_red, label = species),
    size = 4.4,
    fontface = "italic",
    color = "grey40",
    nudge_x = -0.04,      # Pushes all labels 0.2 units to the LEFT
    direction = "y",     # Once pushed left, stack them vertically (prevents messy angles)
    hjust = 1,           # Right-align the text so it looks neat pointing to the data
    xlim = c(-Inf, 0.05),
    box.padding = 0.5,
    max.overlaps = 50,
    segment.alpha = 0.5) +
  
  # F. Styling
  scale_color_manual(name = "Population", 
                     values = c("Native Range (Red Sea)" = "#D9643E", 
                                "Invaded Range (Med)" = "#F8B954")) +
  scale_fill_manual(name = "Population", 
                    values = c("Native Range (Red Sea)" = "#D9643E", 
                               "Invaded Range (Med)" = "#F8B954")) +
  
  theme_classic() +
  
  labs(
    x = "PCoA 1",
    y = "PCoA 2",
    title = "B) Habitat shifts"#,
    #caption = stats_caption
  ) +
  coord_cartesian(xlim = c(-0.1787702, 0.2353858),
                  ylim = c(-0.14, 0.21))

print(shared_sp_plot)
```

##Bind
```{r}
# --- 3. COMBINE & SAVE ---
# Arrange side by side (ncol = 2)
combined_plot <-
  grid.arrange(
    plot_hulls_all+
      theme(
    legend.position = c(0.22, 0.85), 
    legend.background = element_blank(),
    legend.key = element_blank(),
    legend.key.size = unit(0.75, "cm"),
    legend.title = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 12),
    legend.margin = margin(0,0,0,0),
    plot.caption = element_text(hjust = 0.5, face = "italic", size =11),
    plot.caption.position = "plot", # Ensures it aligns to the image edge, not just the panel
    axis.title = element_blank(),
    title = element_text(size=14)
  ),
    shared_sp_plot+
    theme(
    plot.caption = element_text(hjust = .5, face = "italic", size =11),
    plot.caption.position = "plot",
    axis.title = element_blank(),
    title = element_text(size=14)
  ),
    nrow = 2,
    left = "PCoA 2",
    bottom = "PCoA 1"
  )

# Save
ggsave("./plots/Figure_3.png", combined_plot,
       width = 20, height = 30, units = "cm", dpi = 300,bg = "white")
ggsave("./plots/Figure_3.pdf", combined_plot,
       width = 20, height = 30, units = "cm", dpi = 300,bg = "white")
```
